---
title: "Build your config from examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{config-examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(rstudiovim)
```

First it's nice to know you can have a very quick start up

Then we'll talk through some of examples to give you some ideas which you can pick,
ignore and tweak as you please.

## Build from a template

A few simple examples are provided in the 
[basic.vimrc](https://github.com/davidfoord1/rstudiovim/blob/main/inst/basic.vimrc).
You can see the example at the `rsvim_example_path()` and you can use it to start
building your own config with:

```r
rsvim_use_template()
```

This function takes the basic examples, writes it to a `.virmc` file and opens
it for you to make your own edits. You can also specify the files to copy to and
from:

```r
rsvim_use_template(
  to = "path/to/your/.virmc",
  from = "path/to/another/template/.vimrc"
)
```

## Basic Example

<details>
<summary>See the full example config</summary>
```{r comment="", collapse=FALSE}
writeLines(readLines(rsvim_example_path()))
```
</details>

1. Home row exit from insert mode - `inoremap jk <Esc>`

You might have seen enough of this example by now, but we're covering it anyway.
In `i` insert mode, we non-recursively `noremap` map the key chain `jk` to the
key chain `<Esc>`. `Esc` is a standard way to exit from insert to normal mode,
so this gives as a home row way to do the same. That can be pretty neat, by do
consider whether you're likely to be typing `jk` while in insert mode. We use
non-recursive mapping for safety, so for instance if you map `Esc` to something
else, `jk` will still have the expected default `Esc` behaviour. Note there iNULLs
also the `Ctrl+[` method for leaving insert mode, that might be sufficiently
convenient compared to `Esc`.

2. Multiple replace paste - `noremap \p "0p` 

3. New line in normal mode 

`noremap \o o<Esc>`
`noremap \O O<Esc>`

## Full example

A lot more examples are given in the full example, which you can build from with

```r
rsvim_use_template(from = rsvim_example_path("full_example"))
```
